\LaTeX\ Author Guidelines for ICCV Proceedings

### 2019-05-19 chapters/0_abstract.tex
Finish polishing intro, abstract, conclusions

% !TEX root = ../agglo_clust_review.tex
%%%%%%%%% ABSTRACT
\begin{abstract}
We propose a novel theoretical framework that generalizes algorithms for hierarchical agglomerative clustering to weighted graphs with both attractive and repulsive interactions between the nodes. This framework defines an underlying basic algorithm and allows us to explore its combinations with different linkage criteria and cannot-link constraints, i.e. mutual exclusion between clusters. These different combinations then correspond to existing clustering methods as well as new algorithms that we introduce and analyze. We evaluate these different algorithms on instance segmentation, which is the task of partitioning an image into distinct segments, by predicting edge weights with a CNN. \UPDATE{Summarize results} 
\end{abstract}

>>>>

% !TEX root = ../agglo_clust_review.tex
%%%%%%%%% ABSTRACT
\begin{abstract}
We propose a novel theoretical framework that generalizes algorithms for hierarchical agglomerative clustering to weighted graphs with both attractive and repulsive interactions between the nodes. This framework defines an underlying basic algorithm and allows us to explore its combinations with different linkage criteria and cannot-link constraints, i.e. mutual exclusion between clusters. We then prove how these different combinations correspond to existing clustering methods as well as new algorithms that we introduce and analyze. 
An extensive comparison study is performed to evaluate properties of the clustering algorithms, e.g. robustness and efficiency, by applying them to instance segmentation, which is the task of partitioning an image into distinct segments. 
% and by predicting edge weights with a CNN. 
We then show how one of the new algorithms proposed in our framework outperforms all previously known agglomerative methods for signed graphs, both on the competitive CREMI 2016 EM segmentation benchmark and the CityScapes dataset. 
\end{abstract}

### 2019-05-19 chapters/1_intro.tex
Finish polishing intro, abstract, conclusions

\section{Introduction}
The clustering of weighted graphs has been successfully applied to such tasks as image segmentation, tracking... \UPDATE{REFS}
Most graph clustering methods work with positive edge weights only, representing similarities or distances between the nodes. These methods are parameter-based and require users to specify the desired numbers of clusters or a termination criterion (e.g. spectral clustering or iterated normalized cuts) or even a stronger version of supervision in terms of seeds (e.g. seeded watershed or random walker).  
Hierarchical clustering (HC) is a popular graph clustering method, which creates a hierarchy of clusters. Agglomerative HC is a bottom-up approach starting with each node assigned to its own cluster and incrementally merging clusters while moving up the hierarchy \cite{lance1967general}. This method usually requires the user to choose a level in the cluster hierarchy defining the desired output clustering.

Other graph clustering methods work with so-called \emph{signed graphs}, which include both positive and negative edge weights corresponding to attraction and repulsion between nodes. The advantage of using signed graphs over positive-weight graphs is that balancing attraction and repulsion allows us to perform the clustering without defining additional parameters. This can be done optimally by solving the so-called \emph{multicut optimization} or \emph{correlation clustering} problem \cite{kappes2011globally,chopra1991multiway}. However, this problem is NP-hard, so instead of finding optimal solutions, most of the methods perform approximations. Greedy agglomerative clustering algorithms are among the most efficient methods \cite{keuper2015efficient,levinkov2017comparative,wolf2018mutex,kardoostsolving}. 
% Some of these methods \cite{levinkov2017comparative,wolf2018mutex} dynamically introduce cannot-link constraints, representing mutual exclusion relationships between clusters.

Thus, greedy agglomerative clustering algorithms for signed graphs have clear advantages: they are parameter free and efficient. Despite the fact that there exist a variety of these algorithms, there has been no overarching studies comparing them and making it possible to choose the most appropriate algorithm for particular applications and assess their properties, e.g. robustness and efficiency.

@@

In this paper, we propose a novel theoretical framework for generalizing over the greedy agglomerative clustering algorithms for signed graphs by linking these algorithms to the agglomerative HC. This framework defines an underlying basic algorithm and allows us to explore its combinations with different linkage criteria and \emph{cannot-link constraints}, i.e. mutual exclusion between clusters, in a consistent way. These different combinations then correspond to existing clustering algorithms as well as allows us to introduce and explore new algorithms. The framework can also be easily expended to accommodate additional linkage criteria.

%As a first contribution, we reformulate several of these recently proposed graph agglomerative algorithms in a simple review framework (Sec. \ref{sec:general_framework}) that represents a generalization of agglomerative HC to graphs with both attractive and repulsive edge weights (see Figure \ref{fig:intro_figure} for a visual abstract). From this new point of view, agglomerative HC and algorithms proposed in \cite{levinkov2017comparative,wolf2018mutex,lance1967general,kardoostsolving} all share a general procedure but differ in the way inter-cluster interaction is updated after each iteration (Table \ref{tab:linkage-criteria}). At the same time, we also show that this framework allows us to introduce totally new variations of these algorithms. 
%Ultimately, by presenting this simple general approach, our hope is to motivate a deeper understanding of the connection between algorithms for unsigned and signed graphs.

We then evaluate and compare these algorithms on \emph{instance segmentation}, which is a computer vision task consisting in assigning each pixel of an image to an object instance. %, where the number of instances is usually not known in advance. 
We use a CNN to predict the edge weights of a graph such that each node represents a pixel of the image, similarly to \cite{liu2018affinity,lee2017superhuman,wolf2018mutex}, and provide these weights as input to the algorithms in our framework.
% For our experiments, we focus on a \emph{proposal-free} deep learning method that does not involve object detection but directly group pixels into object instances. 
Recently, there has been a growing interest for such \emph{proposal-free} methods that do not involve object detection, since in certain types of data objects cannot be approximated by bounding boxes.
%  (for example in electron microscopy (EM) image volumes of neurons \cite{arganda2015crowdsourcing})

By applying this method both to 2D urban scenes from the CityScapes dataset and the challenging neuron-segmentation task of 3D EM image stacks, \UPDATE{we show how one of the less-known algorithms included in our framework consistently outperformed all other previously proposed signed graph clustering methods and achieved comparable/superior scores to other complex multi-step pipelines...}
% Most recently, success in instance segmentation (IS) has been achieved by applying deep learning and there are two main types of successful deep learning approaches to IS: proposal-based \cite{he2017mask,dai2016instance,li2017fully} and proposal-free \cite{kong2018recurrent,novotny2018semi,kulikov2018instance,kirillov2017instancecut} methods. Proposal-based methods consist of two steps: detecting objects, for example by finding bounding boxes, and assigning pixels to the detected instances. Although these approaches have proven to be highly successful in instance segmentation competitions like MS COCO \cite{lin2014microsoft}, Pascal VOC2012 \cite{everingham2010pascal} and CityScapes \cite{cordts2016cityscapes}, they are not applicable to certain types of data, for example electron microscopy (EM) image volumes of neurons \cite{arganda2015crowdsourcing}, where objects are not approximated well by bounding boxes. 
% More motivation: They are at the same time limited by the quality of the object detection routine, which is hard to train on small datasets (but no ref for this)
% Thus, there is a growing interest for more efficient proposal-free methods that perform IS by directly predicting \UPDATE{pixel features} and then grouping pixels into object instances. In this work, we focus on a common proposal-free method, where a Convolutional Neural Network (CNN) predicts affinities representing how likely it is for pairs of pixels to belong to the same instance. Recently, this approach was successfully applied both to neuron segmentation of 3D EM image stacks \cite{lee2017superhuman,wolf2018mutex} and instance segmentation of 2D urban scenes \cite{liu2018affinity}, where it achieved performances superior and comparable, respectively, to proposal-based methods.
% In this approach, the output of the CNN can be represented as a weighted grid graph such that each node represents a pixel of the image and the weights of the edges define interactions between the pixels. A graph clustering algorithm is then applied to cluster pixels into instances. 
% Finally, the algorithms included in the framework are evaluated on an instance segmentation task, both on 3D biological image stacks (Sec. \ref{sec:neuro_segm_exp}) and 2D urban scenes (Sec. \ref{sec:cityscapes_exp}). Our analysis highlights strengths and weaknesses of each algorithm by comparing their robustness to noise, efficiency and inclination to output a well balanced graph partitioning.

\begin{figure}[t]
\centering
% \includegraphics[width=0.5\textwidth,trim=0.4in 1.2in 0.in 0.05in,clip]{./figs/intro_image.jpg} % left bottom right top
\includegraphics[width=\textwidth]{./figs/intro_image.pdf} % left bottom right top
\caption{
 (\textbf{a}) Iterations of three different algorithms in the generalized framework on  toy graph examples with attractive (green) and repulsive (red) interactions. (\textbf{b}) Example of CREMI 2016 neuron-segmentation data \cite{cremiChallenge} overlaid with instance segmentation. (\textbf{c}) Agglomeration order, representing which pairs of neighboring pixels were merged first (white), later on (brown), or never (black).
%We note how similar linkage criteria, like \emph{Sum} (in the middle) and \emph{Average} (on the right), lead to strongly different merging orders. 
% On the left: a single cluster is returned with only attractive interactions.
 % main ideas and contributions
\label{fig:intro_figure}}
\end{figure}

>>>>

\section{Introduction}
\UPDATE{\today,\currenttime} -  In computer vision, the clustering of weighted graphs has been successfully applied to such tasks as image segmentation, object tracking and pose estimation. %image segmentation, object tracking, stereo matching, etc.), to high-level tasks (e.g. image classification, object recognition, image parsing
% to model many types of re- lations and processes in physical, biological, social and information systems
Most graph clustering methods work with positive edge weights only, which can be interpreted as similarities or distances between the nodes. These methods are parameter-based and require users to specify the desired numbers of clusters or a termination criterion (e.g. spectral clustering or iterated normalized cuts) or even a supervision in terms of seeds (e.g. seeded watershed or random walker).  
%Hierarchical clustering is a popular graph clustering method, which creates a hierarchy of clusters. Hierarchical agglomerative clustering (HAC) is a bottom-up approach starting with each node assigned to its own cluster and incrementally merging clusters while moving up the hierarchy \cite{lance1967general}. This method usually requires the user to choose a level in the cluster hierarchy defining the desired output clustering.

Other graph clustering methods work with so-called \emph{signed graphs}, which include both positive and negative edge weights corresponding to attraction and repulsion between nodes. The advantage of using signed graphs over positive-weighted graphs is that balancing attraction and repulsion allows us to perform the clustering without defining additional parameters. This can be done optimally by solving the so-called \emph{multicut optimization} or \emph{correlation clustering} problem \cite{kappes2011globally,chopra1991multiway}. However, this problem is NP-hard, so instead of finding optimal solutions, several greedy agglomerative clustering algorithms were proposed. 
% Some of these methods \cite{levinkov2017comparative,wolf2018mutex} dynamically introduce cannot-link constraints, representing mutual exclusion relationships between clusters.

Thus, agglomerative clustering algorithms for signed graphs have clear advantages: they are parameter free and efficient. Despite the fact that there exist a variety of these algorithms \cite{keuper2015efficient,levinkov2017comparative,wolf2018mutex,kardoostsolving}, there has been no overarching studies comparing them and making it possible to choose the most appropriate algorithm for particular applications and assess their properties, e.g. robustness and efficiency.

@@

In this paper, we propose a novel theoretical framework for generalizing over agglomerative algorithms for signed graphs by linking them to hierarchical agglomerative clustering, which is a popular bottom-up approach defining an hierarchy of clusters on positive-weighted graphs \cite{lance1967general}. This framework defines an underlying basic algorithm and allows us to explore in a consistent way its combinations with different linkage criteria and \emph{cannot-link constraints}, i.e. mutual exclusion between clusters. We then theoretically prove that different combinations correspond to existing clustering algorithms as well as new algorithms that we introduce and explore. % The framework can also be easily expended to accommodate additional linkage criteria.
%As a first contribution, we reformulate several of these recently proposed graph agglomerative algorithms in a simple review framework (Sec. \ref{sec:general_framework}) that represents a generalization of agglomerative HC to graphs with both attractive and repulsive edge weights (see Figure \ref{fig:intro_figure} for a visual abstract). From this new point of view, agglomerative HC and algorithms proposed in \cite{levinkov2017comparative,wolf2018mutex,lance1967general,kardoostsolving} all share a general procedure but differ in the way inter-cluster interaction is updated after each iteration (Table \ref{tab:linkage-criteria}). At the same time, we also show that this framework allows us to introduce totally new variations of these algorithms. 
%Ultimately, by presenting this simple general approach, our hope is to motivate a deeper understanding of the connection between algorithms for unsigned and signed graphs.

We evaluate and compare these algorithms on \emph{instance segmentation}, which is a computer vision task consisting in assigning each pixel of an image to an object instance. %, where the number of instances is usually not known in advance. 
We use a CNN to predict the edge weights of a graph such that each node represents a pixel of the image, similarly to \cite{liu2018affinity,lee2017superhuman,wolf2018mutex}, and provide these weights as input to the algorithms in our framework.
% For our experiments, we focus on a \emph{proposal-free} deep learning method that does not involve object detection but directly group pixels into object instances. 
%  (for example in electron microscopy (EM) image volumes of neurons \cite{arganda2015crowdsourcing})

With our comparison experiments, performed both on 2D urban scenes and 3D electron microscopy image volumes of neurons, we evaluate the properties of the algorithms in our framework, focusing on their efficiency, robustness and tendency to over- or under-cluster.
Our findings show that one of the new agglomerative algorithms introduced by our generalized framework, based on an average linkage criterion, consistently outperforms all other previously known agglomerative algorithms for signed graphs and achieves competitive scores on the challenging CREMI 2016 segmentation benchmark.
% This simple agglomerative algorithm also achieves competitive scores on the challenging CREMI 2016 segmentation benchmark, similarly to other approaches that rely on complex multi-step pipelines.

\begin{figure}[t]
\centering
% \includegraphics[width=0.5\textwidth,trim=0.4in 1.2in 0.in 0.05in,clip]{./figs/intro_image.jpg} % left bottom right top
\includegraphics[width=\textwidth]{./figs/intro_image.pdf} % left bottom right top
\caption{
 (\textbf{a}) Iterations of three different algorithms in the generalized framework on  toy graph examples with attractive (green) and repulsive (red) interactions. (\textbf{b}) Example of CREMI 2016 neuron-segmentation data \cite{cremiChallenge} overlaid with instance segmentation. (\textbf{c}) Agglomeration order, representing which pairs of neighboring pixels were merged first (white), later on (brown), or never (black).
 \TODO{re-arrange, show affinities and replace unsigned case with Abs Max}
%We note how similar linkage criteria, like \emph{Sum} (in the middle) and \emph{Average} (on the right), lead to strongly different merging orders. 
% On the left: a single cluster is returned with only attractive interactions.
 % main ideas and contributions
\label{fig:intro_figure}}
\end{figure}

### 2019-05-19 chapters/3_algorithm.tex
Finish polishing intro, abstract, conclusions

\section{Generalized framework for Signed Graph Agglomerative Clustering} \label{sec:general_framework}
In this section, we first define the graph notation in Sec.~\ref{sec:notation} and then introduce one of our main contributions: a signed graph partitioning algorithm (Sec. \ref{sec:algorithm}) that can be seen as a generalization of several existing and new clustering algorithms (Sec. \ref{sec:alg_update_rules}).
 % introduces the proposed generalized framework: 
%that represents a simple \UPDATE{generalized} formalization of many agglomerative graph clustering algorithms. \UPDATE{It can be used to describe both unsigned clustering algorithms ingesting positive node similarities and signed clustering algorithms using attractive and repulsive cues.}\\
% we  in Sec.~\ref{sec:notation}, introduce Algorithm \ref{main_alg} in  and finally show how it can be seen as a generalization of existing and new graph clustering algorithms in .

@@

We consider an undirected simple edge-weighted graph $\mathcal{G}(V,E,w^+, w^-)$ with both attractive and repulsive edge attributes, in which nodes can represent either pixels, superpixels or voxels. We call the set $\Pi$ a \emph{clustering} or \emph{partitioning} with $K$ clusters if $V = \cup_{i=1}^K S_i $, $\,S_i \cap S_j = \emptyset$ for each $i\neq j$ and every $S \in \Pi$ induces a connected subgraph of $\mathcal{G}$. We also denote as $S_u$ the cluster associated with node $u$.
% \begin{equation}
% S_u \equiv S \in \Pi \,\, \text{s.t.} \,\, u \in S
% \end{equation}
% In this work, we consider the problem of clustering a weighted graph $\mathcal{G}(V,E,w^+, w^-)$ with both attractive and repulsive edge attributes. 
The weight function $w^+: E \rightarrow \mathbb{R}^+$ associates a positive scalar attribute $w_e^+\in \mathbb{R}^+$ representing a merge affinity or a similarity measure: the higher this number, the higher the inclination of the two incident vertices to be assigned to the same cluster\footnote{Note that other formalisms \UPDATE{REFS} for positively weighted graphs associate distances to the edges, thus, the \emph{lower} the edge weight, the higher the attraction between the two linked nodes, contrary to our definition of $w^+$.}. On the other hand, $w^-: E \rightarrow \mathbb{R}^+$ associates a split tendency $w_e^- \in \mathbb{R}^+$: the higher this weight, the more the incident vertices would like to be in different clusters. 
Graphs of the type $\mathcal{G}(V,E,w^+, w^-)$ are also often defined as \emph{signed graphs} $\mathcal{G}(V,E,\cost)$, featuring positive and negative edge weights $\cost_e\in \mathbb{R}$. Following the theoretical considerations in \cite{lange2018partial}, we define these signed weights as ${\cost_e = w_e^+ - w_e^-}$. Some approaches directly compute $\cost_e$, whereas others compute $w_e^+$ and $w_e^-$ separately.
In this formalism, graphs with purely attractive interactions are a special case of $\mathcal{G}(V,E,\cost)$ with $\cost_e \geq 0, \, \forall e \in E$.

\textbf{Inter-cluster interaction } We call two clusters $S_1,S_2$ \emph{adjacent} if there exists at least one edge ${e_{uv}\in E}$ connecting a node $u\in S_1$ to a node $v\in S_2$. In agglomerative HC, the interaction $\interact(S_1,S_2)$ between the two clusters is usually defined as a function, named \emph{linkage criterion}, depending on the weights of \emph{all} edges connecting clusters $S_1$ and $S_2$. 
All the linkage criteria tested in this article are listed and defined in Table \ref{tab:linkage-criteria}.
% For example, for a \emph{maximum} linkage criterion statistic (single-linkage), the inter-cluster interaction is given by:
% \begin{equation}\label{eq:max_linkage}
% \interact_{S_1,S_2} = \, \max_{e\in U(S_u,S_v)} \cost(e), \qquad \text{where}\quad E_{12} = \{ e_{uv} \in E | u \in S_1, \, v \in S_2 \}.
% \end{equation}
% Other common choices are \emph{minimum} (complete linkage) or \emph{average} (average linkage). %and \emph{sum}.

\begin{figure}
\centering
        \begin{subfigure}[t]{0.46 \textwidth}
        \centering
        \includegraphics[width=\textwidth]{./figs/example_no_constr.pdf}
        \caption{No constraints}\label{subfig:no_constraints}
    \end{subfigure} \hfill
    \begin{subfigure}[t]{0.46 \textwidth}
        \centering
        \includegraphics[width=\textwidth]{./figs/example_with_constr.pdf}
        \caption{With constraints}\label{subfig:with_constraints}
    \end{subfigure}
\caption{Some iterations of the generalized algorithm (using \emph{Sum} linkage criteria) with and without adding cannot-link constraints. The graph has both attractive (green) and repulsive (red) edges and cannot-link constraints are shown with triple violet bars on the edges. We note that when constraints are enforced, the final clustering is given by two clusters instead of only one.}
\label{fig:algorithm_with_without_CLC}
\end{figure}
\begin{algorithm}[t]
  \caption{\algname{}: Generalized Algorithm for Agglomerative Signed Graph Partitioning}
   \hspace*{\algorithmicindent} \textbf{Input:} Graph $\mathcal{G}(V,E,w^+,w^-)$; linkage criterion $\interact{}$; boolean {\color{blue}\texttt{addCannotLinkConstraints}}  \\
  \hspace*{\algorithmicindent} \textbf{Output:} Final clustering $\Pi$\\
  \hspace*{\algorithmicindent} 
  \begin{algorithmic}[1]
      \State Initialize clustering $\Pi=\{\{v_1\}, \ldots, \{v_N\}\}$ with each node in its own cluster
      \State Initial interactions between nodes given by $\cost_e = w^+_e - w^-_e$
      % \State Initialize \texttt{canBeMerged}$[e] \gets$ \texttt{True} $\,\,\, \forall e\in E$
      % \State Sort edges in priority queue (PQ) in descending order of absolute weight $|\cost_e|=|w^+_e - w^-_e|$ 
      % \State
      \Repeat
        \State Select pair of clusters $S_u,S_v\in\Pi$ with highest absolute interaction $|\interact{}(S_u,S_v)|$
        % \State
        \If{\big[{\color{ForestGreen}\textbf{$\interact{}(S_u,S_v) > 0$}}\big] \textbf{and} \big[$S_u,S_v$ are \textbf{not} constrained\big]}
          \State Merge cluster $S_u$ with $S_v$: update interactions and cannot-link constraints with all their neighbors
        % \State
        \ElsIf{\big[{\color{red}\textbf{$\interact{}(S_u,S_v) \leq 0$}}\big] \textbf{and} {\color{blue}\texttt{addCannotLinkConstraints}}}
          \State Add CannotLink Constraint between clusters $S_u$ and $S_v$
        \EndIf
      \Until{\big[all interactions between clusters are repulsive\big] \textbf{or} \big[all adjacent clusters are constrained\big]}
        % \If{({\color{ForestGreen}\textbf{$\interact{}(S_u,S_v) > 0$}}) \textbf{and} \texttt{canBeMerged}$[e_{uv}]$}
        %   \State Contract $e_{uv}$ in $\tilde{\mathcal{G}}_\Pi$ and merge the two clusters $S_u,S_v \in \Pi$
        %   \For{every new pair of double edges $(e_1,e_2)$ in $\tilde{\mathcal{G}}_\Pi$}
        %     % \State Get costs $\cost_1, \cost_2$ of $e_1,e_2$
        %     \State Delete $e_2$ from $\tilde{\mathcal{G}}_\Pi$ and update priority of $e_1$ with rule $f(\tilde{\cost}_{e_1},\tilde{\cost}_{e_2})$ defined in Tab. \ref{tab:linkage-criteria} 
        %   \EndFor
        % \EndIf
        % \If{({\color{red}\textbf{$\tilde{\cost} \leq 0$}}) \textbf{and} {\color{blue}\texttt{addCannotLink}}}
        %   \State \texttt{canBeMerged}$[e_{uv}] \gets$ \texttt{False}
        % \EndIf

@@

        \end{tabular}
        % \caption{Linkage criteria}
    \end{subtable} 
    \caption{The table lists the existing clustering algorithms that can be reformulated as special cases of the proposed generalized algorithm \algname{}, given a linkage criteria, a type of graph (signed or unsigned) and the optional use of cannot-link constraints. The set $E_{uv}$ is defined as the set of all edges connecting cluster $S_u$ to cluster $S_v$, i.e. $E_{uv}=(S_u \times S_v) \cap E$.}
    \label{tab:linkage-criteria}
\end{table*}
\begin{figure}
\centering
% \includegraphics[width=0.5\textwidth,trim=0.4in 1.2in 0.in 0.05in,clip]{./figs/intro_image.jpg} % left bottom right top
\includegraphics[width=\textwidth]{./figs/comparison.pdf} % left bottom right top
\caption{Failure cases of \algname{} with different linkage criteria highlighted on some difficult parts of the CREMI Challenge \cite{cremiChallenge} training data. Only the few wrongly segmented regions are highlighted in colors. Red arrows point to wrongly split regions. Yellow arrows point out merge errors. Green arrows indicate challenging, but correctly segmented regions. 
% Only the involved segments are colored.
 % main ideas and contributions
\label{fig:cremi_comparison}}
\end{figure}

@@

Given a graph with both attractive and repulsive cues, an edge contraction algorithm with a sum update rule was already proposed in \cite{levinkov2017comparative,keuper2015efficient} (Table \ref{tab:linkage-criteria}, \emph{Sum} linkage). They present both a version with cannot-link constraints and one without, and then compare them with other greedy local-search algorithms solving the multicut optimization problem.
The Mutex Watershed \cite{wolf2018mutex} is another signed graph partitioning algorithm that introduces dynamical cannot-link constraints. In Appendix \ref{sec:equivalence_MWS} we prove that it can be seen as an efficient implementation of \algname{} with \emph{Absolute maximum} linkage (see def. in Table \ref{tab:linkage-criteria}) and observe that in this case \algname{} returns the same clustering with or without enforcing cannot-link constraints.
\UPDATE{Less-known options: \emph{Average}, \emph{Max} or \emph{Min} linkage applied to signed graphs or never combined with cannot-link constraints.}

>>>>

\section{Generalized Framework for Agglomerative Clustering of Signed Graphs} \label{sec:general_framework}
In this section, we first define notation and then introduce one of our main contributions: a signed graph partitioning algorithm (Sec. \ref{sec:algorithm}) that can be seen as a generalization of several existing and new clustering algorithms (Sec. \ref{sec:alg_update_rules}).
 % introduces the proposed generalized framework: 
%that represents a simple \UPDATE{generalized} formalization of many agglomerative graph clustering algorithms. \UPDATE{It can be used to describe both unsigned clustering algorithms ingesting positive node similarities and signed clustering algorithms using attractive and repulsive cues.}\\
% we  in Sec.~\ref{sec:notation}, introduce Algorithm \ref{main_alg} in  and finally show how it can be seen as a generalization of existing and new graph clustering algorithms in .

@@

We consider an undirected simple edge-weighted graph $\mathcal{G}(V,E,w^+, w^-)$ with both attractive and repulsive edge attributes. In computer vision applications, the nodes can represent either pixels, superpixels or voxels. We call the set $\Pi$ a \emph{clustering} or \emph{partitioning} with $K$ clusters if $V = \cup_{i=1}^K S_i $, $\,S_i \cap S_j = \emptyset$ for each $i\neq j$ and every $S \in \Pi$ induces a connected subgraph of $\mathcal{G}$. We also denote as $S_u$ the cluster associated with node $u$.
% \begin{equation}
% S_u \equiv S \in \Pi \,\, \text{s.t.} \,\, u \in S
% \end{equation}
% In this work, we consider the problem of clustering a weighted graph $\mathcal{G}(V,E,w^+, w^-)$ with both attractive and repulsive edge attributes. 
The weight function $w^+: E \rightarrow \mathbb{R}^+$ associates a positive scalar attribute $w_e^+\in \mathbb{R}^+$ representing a merge affinity or a similarity measure: the higher this number, the higher the inclination of the two incident vertices to be assigned to the same cluster\footnote{Note that other formalisms \UPDATE{REFS} for positively weighted graphs associate distances to the edges, thus, the \emph{lower} the edge weight, the higher the attraction between the two linked nodes, contrary to our definition of $w^+$.}. On the other hand, $w^-: E \rightarrow \mathbb{R}^+$ associates a split tendency $w_e^- \in \mathbb{R}^+$: the higher this weight, the more the incident vertices would like to be in different clusters. 
Graphs of the type $\mathcal{G}(V,E,w^+, w^-)$ are also often defined as \emph{signed graphs} $\mathcal{G}(V,E,\cost)$, featuring positive and negative edge weights $\cost_e\in \mathbb{R}$. Following the theoretical considerations in \cite{lange2018partial}, we define these signed weights as ${\cost_e = w_e^+ - w_e^-}$. Some approaches directly compute $\cost_e$, whereas others compute $w_e^+$ and $w_e^-$ separately.
In this formalism, graphs with purely attractive interactions are a special case of $\mathcal{G}(V,E,\cost)$ with $\cost_e \geq 0, \, \forall e \in E$.

\textbf{Inter-cluster interaction } We call two clusters $S_i,S_j$ \emph{adjacent} if there exists at least one edge ${e_{uv}\in E}$ connecting a node $u\in S_i$ to a node $v\in S_j$. In agglomerative HC, the interaction $\interact(S_i,S_j)$ between the two clusters is usually defined as a function, named \emph{linkage criterion}, depending on the weights of \emph{all} edges connecting clusters $S_i$ and $S_j$. 
All the linkage criteria tested in this article are listed and defined in Table \ref{tab:linkage-criteria}.
% For example, for a \emph{maximum} linkage criterion statistic (single-linkage), the inter-cluster interaction is given by:
% \begin{equation}\label{eq:max_linkage}
% \interact_{S_1,S_2} = \, \max_{e\in U(S_u,S_v)} \cost(e), \qquad \text{where}\quad E_{12} = \{ e_{uv} \in E | u \in S_1, \, v \in S_2 \}.
% \end{equation}
% Other common choices are \emph{minimum} (complete linkage) or \emph{average} (average linkage). %and \emph{sum}.

\begin{figure}
\centering
        \begin{subfigure}[t]{0.46 \textwidth}
        \centering
        \includegraphics[width=\textwidth]{./figs/example_no_constr.pdf}
        \caption{No constraints}\label{subfig:no_constraints}
    \end{subfigure} \hfill
    \begin{subfigure}[t]{0.46 \textwidth}
        \centering
        \includegraphics[width=\textwidth]{./figs/example_with_constr.pdf}
        \caption{With cannot-link constraints}\label{subfig:with_constraints}
    \end{subfigure}
\caption{Some iterations of the generalized algorithm (using \emph{Sum} linkage criteria) with and without adding cannot-link constraints. The graph has both attractive (green) and repulsive (red) edges and cannot-link constraints are shown with triple violet bars on the edges. We note that when constraints are enforced, the final clustering is given by two clusters instead of only one.}
\label{fig:algorithm_with_without_CLC}
\end{figure}
\begin{algorithm}[t]
  \caption{\algname{}: Generalized Algorithm for Agglomerative Signed Graph Partitioning}
   \hspace*{\algorithmicindent} \textbf{Input:} Graph $\mathcal{G}(V,E,w^+,w^-)$; linkage criterion $\interact{}$; boolean {\color{blue}\texttt{addCannotLinkConstraints}}  \\
  \hspace*{\algorithmicindent} \textbf{Output:} Final clustering $\Pi$\\
  \hspace*{\algorithmicindent} 
  \begin{algorithmic}[1]
      \State Initialize clustering $\Pi=\{\{v_1\}, \ldots, \{v_N\}\}$ with each node in its own cluster
      \State Initial interactions between nodes given by $\cost_e = w^+_e - w^-_e$
      % \State Initialize \texttt{canBeMerged}$[e] \gets$ \texttt{True} $\,\,\, \forall e\in E$
      % \State Sort edges in priority queue (PQ) in descending order of absolute weight $|\cost_e|=|w^+_e - w^-_e|$ 
      % \State
      \Repeat
        \State Select pair of clusters $S_u,S_v\in\Pi$ with highest absolute interaction $|\interact{}(S_u,S_v)|$
        % \State
        \If{\big[{\color{ForestGreen}\textbf{$\interact{}(S_u,S_v) > 0$}}\big] \textbf{and} \big[$S_u,S_v$ are \textbf{not} constrained\big]}
          \State Merge cluster $S_u$ with $S_v$: update interactions and cannot-link constraints with all their neighbors
        % \State
        \ElsIf{\big[{\color{red}\textbf{$\interact{}(S_u,S_v) \leq 0$}}\big] \textbf{and} {\color{blue}\texttt{addCannotLinkConstraints}}}
          \State Add CannotLink Constraint between clusters $S_u$ and $S_v$
        \EndIf
      \Until{\big[all interactions between clusters are repulsive\big] \textbf{or} \big[all adjacent clusters have cannot-link constraints\big]}
        % \If{({\color{ForestGreen}\textbf{$\interact{}(S_u,S_v) > 0$}}) \textbf{and} \texttt{canBeMerged}$[e_{uv}]$}
        %   \State Contract $e_{uv}$ in $\tilde{\mathcal{G}}_\Pi$ and merge the two clusters $S_u,S_v \in \Pi$
        %   \For{every new pair of double edges $(e_1,e_2)$ in $\tilde{\mathcal{G}}_\Pi$}
        %     % \State Get costs $\cost_1, \cost_2$ of $e_1,e_2$
        %     \State Delete $e_2$ from $\tilde{\mathcal{G}}_\Pi$ and update priority of $e_1$ with rule $f(\tilde{\cost}_{e_1},\tilde{\cost}_{e_2})$ defined in Tab. \ref{tab:linkage-criteria} 
        %   \EndFor
        % \EndIf
        % \If{({\color{red}\textbf{$\tilde{\cost} \leq 0$}}) \textbf{and} {\color{blue}\texttt{addCannotLink}}}
        %   \State \texttt{canBeMerged}$[e_{uv}] \gets$ \texttt{False}
        % \EndIf

@@

        \end{tabular}
        % \caption{Linkage criteria}
    \end{subtable} 
    \caption{The table lists the existing clustering algorithms that can be reformulated as special cases of the proposed generalized algorithm \algname{}, given a linkage criteria, a type of graph (signed or unsigned) and the optional use of cannot-link constraints. The set $E_{uv}$ is defined as the set of all edges connecting cluster $S_u$ to cluster $S_v$, i.e. $E_{uv}=(S_u \times S_v) \cap E$.}
    \label{tab:linkage-criteria}
\end{table*}
\begin{figure}
\centering
% \includegraphics[width=0.5\textwidth,trim=0.4in 1.2in 0.in 0.05in,clip]{./figs/intro_image.jpg} % left bottom right top
\includegraphics[width=\textwidth]{./figs/comparison.pdf} % left bottom right top
\caption{Failure cases of \algname{} with different linkage criteria highlighted on some difficult parts of the CREMI Challenge \cite{cremiChallenge} training data. Only the few \emph{wrongly} segmented regions are highlighted in colors. Note that the data is 3D, hence the same color could be assigned to parts of segments that appear disconnected in 2D.  Red arrows point to wrongly split regions. Yellow arrows point out merge errors. Green arrows indicate challenging, but correctly segmented regions. 
% Only the involved segments are colored.
 % main ideas and contributions
\label{fig:cremi_comparison}}
\end{figure}

@@

Given a graph with both attractive and repulsive cues, an edge contraction algorithm with a sum update rule was already proposed in \cite{levinkov2017comparative,keuper2015efficient} (Table \ref{tab:linkage-criteria}, \emph{Sum} linkage). They present both a version with cannot-link constraints and one without, and then compare them with other greedy local-search algorithms solving the multicut optimization problem.
The Mutex Watershed \cite{wolf2018mutex} is another signed graph partitioning algorithm that introduces dynamical cannot-link constraints. In Appendix \ref{sec:appendix_abs_max} we prove that it can be seen as an efficient implementation of \algname{} with \emph{Absolute maximum} linkage (see def. in Table \ref{tab:linkage-criteria}) and observe that in this case \algname{} returns the same clustering with or without enforcing cannot-link constraints.
On the other hand, to our knowledge, \emph{Average}, \emph{Max} or \emph{Min} linkage criteria have never been used for signed graphs agglomerative algorithms or been combined with cannot-link constraints.
